---
title: "Investigating Social Perceptions of Women’s Rights and Abortion Over Years: A Research with Surveys"
author: 
  - Kuiyao Qiao
  - Xiaoyu Miao
thanks:"Code and data are available at: https://github.com/QPP123/Investigating-Social-Perceptions-of-Women-s-Rights-and-Abortion-Over-Years-A-Research-with-Surveys"
date: March 11, 2024
date-format: long
abstract:"This paper investigates changes in the social perception of abortion in the United States over the past five decades, utilizing data from the General Social Survey conducted by the National Opinion Research Center at the University of Chicago. Through meticulous analysis of survey responses from approximately 46,700 valid cases, we uncover that, although opinions on abortion have slightly shifted, these changes are not statistically significant. Our findings reveal a complex landscape where support for abortion varies across different circumstances, such as the health of the woman, economic constraints, marital status, fetal health issues, and cases of rape, reflecting nuanced societal attitudes towards this contentious issue. This study highlights the critical role of social perception in shaping abortion-related legislation, healthcare access, and the broader discourse on gender equality and women's autonomy, underlining the importance of continuous dialogue and informed policy-making in navigating the complexities of abortion rights."
format: pdf
number-sections: true
ibligrapgy:references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#download the packages if necessary, then load the packages
library(readxl)
library(dplyr)
library(ggplot2)

data <- read_excel("C:/Users/Kuiyao/iCloudDrive/Documents/UT/INF 312/Paper/Paper 3/data/batch_0/result/GSS.xlsx")
```

# Introduction

**The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate;**

**Estimand**

**The estimand is clearly stated in the introduction.**

**1) broader context to motivate:**

Nowadays we live in a world full of dynamic changes 

Women’s rights are especially important. Under this relatively broad topic, abortion is an essential yet controversial subtopic that worth studying, as this relates to lives. The topic is important also because the society’s general perceptions of women’s rights and abortion can greatly affect numerous women’s fates – whether they could be financially self-sufficient or not, whether they get to choose to become a mother or not, whether they could live their life according to their free wills or not. As half of the world’s populations are biologically females, and almost everyone has female family member and/or female friends, the topic is directly related to almost everyone. 

**2) some detail about what the paper is about;**

**3) a clear gap that needs to be filled;**

**We intend to investigate the gap that whether perception of abortion changed significantly over the past 5 decades.**

**4) what was done;**

We gained the data from GSS (General Social Survey), a sociological survey created by the by the National Opnion Research at the University of Chicago. This makes sure the source of data is relatively reliable, and the data is relatively accurate and up-to-date. We downloaded tha survey data and made graphs and table out of them, and then we use those to analyze the patterns. 

**5) what was found;**

We found that over the past 5 years, the social perceptions and opinions of abortion in the United States slightly changed, but the changes were not significant.

**6) why it is important;**

These findings are inportant because social perception on abortion significantly impacts lawmaking, healthcare access, and individual experiences, as public opinion shapes abortion-related legislation, influences the availability of services, and affects the stigmatization or support individuals face. Moreover, these perceptions guide educational campaigns on reproductive health and underscore the connection between abortion views and broader issues of gender equality and women's autonomy. This complex interplay between social attitudes and practical outcomes highlights the importance of understanding and addressing public opinions on abortion. Seeing the historical trends through survey answers in the past 5 decades could tell the information about social perceptions on abortion over decades, and thus lead us to better understand the development of the social movements.

**7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.**

In this paper, we structured it into six parts: abstract, introduction, data, results, discussion, and supplementary survey questions. In each part, we 

**Estimand**: This study aim to estimate the nation’s perception based on the valid survey responses of around 46700 cases. Although the total cases are around 72000, only around 46700 cases are valid. Comparing to the whole population of the United States, this quantity/parameter is not large. However it does represent a general sense of social perceptions. With this being said, we do want to recognize that the limitation of this survey study: that the demography of the respondents might be limited to certain regions/genders/age ranges/socioeconimc statuses, which could impact the inclusiveness and accuracy of the results.

# **Data**

# Results

Results will likely require summary statistics, tables, graphs, images, and possibly statistical analysis or maps. There should also be text associated with all these aspects. Show the reader the results by plotting them where possible. Talk about them. Explain them. That said, this section should strictly relay results. Regression tables must not contain stars.

Graph 1: 

Key points are:

-   The 'Yes' responses (green line) are consistently higher than the 'No' responses (red line) from 1970 to slightly past 2020, indicating a prevailing agreement with the option of abortion under these circumstances.

-   There's a noticeable spike in 'Yes' responses around the year 2000, followed by a sharp decline and then another rise. These fluctuations could be due to a variety of factors, including shifts in public awareness, policy changes, or external events influencing public opinion.

-   The 'No' responses show a slight increase over time but remain relatively low compared to the 'Yes' responses, suggesting that opposition to abortion in the case of potential serious defects is less common.

For this research, these trends could highlight that there is generally strong support for the availability of abortion when there's a risk of serious defects in the baby, but it's important to consider the context behind the fluctuations in responses for a comprehensive analysis.

```{r}
#| echo: false
#| fig-cap: If there is a strong chance of serious defect in the baby?

filtered_data_abdefect <- data %>%
  filter(abdefect %in% c("YES", "NO"))

data_summary_abdefect <- filtered_data_abdefect %>%
  group_by(year, abdefect) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abdefect, aes(x = year, y = count, color = abdefect, group = abdefect)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

Graph 2:  Married women receive abortion?

Observations from the graph:

-   The 'Yes' responses (green) and 'No' responses (red) are quite volatile over the years. There are times when 'Yes' is higher and others when 'No' dominates, suggesting fluctuating opinions on this issue.

-   Around 2000 and just after 2010, there are sharp spikes in both 'Yes' and 'No' responses, which could be tied to specific events or changes in societal attitudes at those times.

-   Despite the fluctuations, 'Yes' responses tend to be higher on average in the later years, especially past 2010, hinting at a possible trend towards greater acceptance of abortion under these circumstances.

In this research, these trends could point to evolving social norms around family planning and autonomy within marriage, with a more recent tilt towards supporting a married woman's choice not to have more children. The variability could also reflect the impact of policy changes or public debates occurring at different times.

```{r}
#| echo: false
#| fig-cap: If she is married and does not want any more children?

filtered_data_abnomore <- data %>%
  filter(abnomore %in% c("YES", "NO"))

data_summary_abnomore <- filtered_data_abnomore %>%
  group_by(year, abnomore) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abnomore, aes(x = year, y = count, color = abnomore, group = abnomore)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

Graph 3: 

Key observations include:

Here's what the data indicates:

-   The 'Yes' responses (green) are substantially higher than the 'No' responses (red) throughout the entire period from the 1970s to 2020. This indicates strong and consistent support for a woman's right to an abortion when her health is at risk.

-   The 'Yes' line shows some fluctuation over the years, with peaks and valleys, but the overall support remains high.

-   The 'No' responses are relatively flat and low in comparison, suggesting that there is minimal opposition to abortion under these circumstances and that this view has not changed dramatically over time.

In the context of this research, this graph could suggest that there is a general consensus supporting the necessity of abortion when it comes to health concerns, and this has been a relatively stable opinion over the last five decades.

```{r}
#| echo: false
#| fig-cap: If the womans own health is seriously endangered by the pregnancy?

filtered_data_abhlth <- data %>%
  filter(abhlth %in% c("YES", "NO"))

data_summary_abhlth <- filtered_data_abhlth %>%
  group_by(year, abhlth) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abhlth, aes(x = year, y = count, color = abhlth, group = abhlth)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

Graph 4:

Key observations are:

-   'Yes' responses (green) generally exceed 'No' responses (red), indicating more support for allowing abortions under these financial circumstances.

-   The data is quite volatile, with many peaks and troughs, suggesting that public opinion on this issue has varied significantly over time.

-   Around the late 1990s and again in the late 2000s, there are moments where 'No' responses surpass 'Yes' responses, which could be indicative of particular events or changes in societal attitudes that influenced public opinion during those times.

This data is valuable for this research as it demonstrates that economic factors affecting decisions on abortion have been a point of contention over the years, with no clear consensus emerging. It might be useful to investigate the societal and economic contexts corresponding to the shifts in public opinion reflected in the graph.

```{r}
#| echo: false
#| fig-cap: If the family has a very low income and cannot afford any more children?

filtered_data_abpoor <- data %>%
  filter(abpoor %in% c("YES", "NO"))

data_summary_abpoor <- filtered_data_abpoor %>%
  group_by(year, abpoor) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abpoor, aes(x = year, y = count, color = abpoor, group = abpoor)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

Graph 5:

Observations include:

-   'Yes' responses (green) are significantly higher than 'No' responses (red) throughout the entire time span, indicating strong support for allowing abortion under these circumstances.

-   There are notable fluctuations in the 'Yes' responses, with peaks and valleys, but the overall trend remains much higher than the 'No' responses.

-   The 'No' responses are relatively consistent over the years and remain low, suggesting that opposition to abortion in this context is consistently minimal.

From this data, one could infer that there is a stable and overwhelming support for a woman's right to choose abortion in the case of rape, which has been maintained over the past five decades. This could be a key point in your research, reflecting a societal consensus on the right to abortion in such traumatic circumstances.

```{r}
#| echo: false
#| fig-cap: If she became pregnant as a result of rape?

filtered_data_abrape <- data %>%
  filter(abrape %in% c("YES", "NO"))

data_summary_abrape <- filtered_data_abrape %>%
  group_by(year, abrape) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abrape, aes(x = year, y = count, color = abrape, group = abrape)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

Graph 6:

Observations include:

-   Unlike the previous graph, here, the 'Yes' (green) and 'No' (red) responses show a more competitive trend, with 'No' responses occasionally matching or exceeding 'Yes' responses at several points in time.

-   The 'Yes' responses show high volatility but maintain an edge over 'No' responses for a majority of the time period.

-   The late 1990s through the 2000s show particularly sharp fluctuations, with 'No' responses peaking above 'Yes' responses more frequently compared to other decades.

This data suggests that public opinion on abortion in the case of rape has been more divided at certain times, with significant changes in attitude that may correlate with cultural, legal, or political events. It could be valuable to correlate these shifts with specific historical contexts for a deeper understanding in your research.

```{r}
#| echo: false
#| fig-cap: If she became pregnant as a result of rape?

filtered_data <- data %>%
  filter(absingle %in% c("YES", "NO"))

data_summary_absingle <- filtered_data %>%
  group_by(year, absingle) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_absingle, aes(x = year, y = count, color = absingle, group = absingle)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

Graph 7:

Notable points:

-   There's a close contention between 'Yes' (green) and 'No' (red) responses throughout the years, reflecting a more divided public opinion on the broader issue of abortion on demand.

-   The 'No' responses have peaks that often surpass the 'Yes' responses, particularly in the mid-1980s to the mid-1990s and again in the early 2000s.

-   Both responses exhibit significant volatility, which could be indicative of shifting social, political, and legal landscapes influencing public opinion on the issue.

This graph suggests that when considering abortion without condition, public opinion has been more polarized, with shifts possibly aligned with broader debates and changes in societal attitudes towards abortion.

```{r}
#| echo: false
#| fig-cap: abortion if woman wants for any reason

data_summary_abany <- data %>%
  filter(abany %in% c("YES", "NO")) %>%
  group_by(year, abany) %>%
  summarise(count = n(), .groups = 'drop') %>%
  mutate(abany = factor(abany, levels = c("YES", "NO"))) # Ensure consistent ordering for the line plot

ggplot(data_summary_abany, aes(x = year, y = count, color = abany, group = abany)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

The seven graphs you provided show public opinion over time on various circumstances under which abortion might be sought. Here's a consolidated summary of insights from the data:

-   Health of the woman: There is consistent and strong support for abortion when the woman's health is seriously endangered. This opinion seems to hold relatively steady over the decades.

-   Economic reasons: Opinions on abortion due to economic reasons, such as the family being of very low income and unable to afford more children, show more fluctuation and suggest that this reason for abortion is more contentious and less consistently supported than health-related reasons.

-   Marital status: There appears to be more support for abortion rights for unmarried women compared to married women who do not want more children, although both show some level of volatility in responses over the years.

-   Fetal health issues: There is strong support for abortion in cases of a strong chance of a serious defect in the baby, but with some fluctuations that might be tied to external factors or events.

-   Pregnancy resulting from rape: Support for abortion is very high when pregnancy is the result of rape, with minimal opposition across the decades.

-   General abortion access: When considering abortion for any reason the woman wants, the data shows a divided opinion with significant contention and volatility over the years.

From these graphs, we can find that public opinion on abortion is generally more supportive when the reasons are related to the health of the woman or the fetus, or in cases of rape. When the reasons for seeking an abortion are not tied to such specific conditions, like economic reasons or general access to abortion for any reason, the public opinion seems to be more divided and subject to change over time.

The volatility in responses, particularly in the graphs related to economic reasons and general abortion access, suggests that these topics are more influenced by the social, political, and economic context at various points in time. These insights could be critical for your research as they demonstrate the complexity and nuance of public opinion on abortion. It would be essential to analyze these trends in relation to policy changes, economic conditions, and social attitudes to understand the underlying causes of the fluctuations in opinion.

# Discussion

Some questions that a good discussion would cover include (each of these would be a sub-section of something like half a page to a page): What is done in this this paper? What is something that we learn about the world? What is another thing that we learn about the world? What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?

What is done in this paper:

This paper examines social perception 

We learn more about women’s rights movement and abortion 

Key Features: 

Strengths: valid cases is large

Weaknesses: 

It does not mention how the data was collected.

audience: targeted to a certain group (age, gender, etc)?

potential bias by methods of collection

What can we proceed in the future: 

Discussion of the questionnaire:

The good part is that each question is very specific and on-point: they are clearly yes or no question but also leave the respondents the option of ‘don’t know’ or ‘not applicable’, which is professionally designed. 

Bad part of the questionnaire is that the questions did not cover all the scenarios: for example, whether or not to receive abortion could be considered very differently if the pregnancy is over 10 or 12 weeks. The survey questions should add questions regarding this, such as ‘do you support abortion if the pregancy is over 10 weeks (which could be medically risky)?’ or something like that.

# Conclusion

# Appendix

**Survey**

**The survey should have an introductory section and include the details of a contact person. The survey questions should be well constructed and appropriate to the task. The questions should have an appropriate ordering. A final section should thank the respondent.**

**Supplementary Survey**

Introduction to the survey: Women’s rights is an essential topic in nowadays society. Within this topic, abortion is a controversial topic. The General Social Survey collects information on people’s opinions and perceptions on those topics. In order to deepen our understandings and analysis of the topics, we designed this survey to collect people’s opinions and perceptions of women’s rights and abortion. We conduct the research studies with good faiths.

Thank you for all the respondents for completing this survey! We appreciate your time and effort. Your response will help us gain a deeper understanding of people’s perceptions on women’s rights and abortion.

Supplementary Survey Questions:

1.  What is your opinion on women’s rights?

    1.  Support it           b. Neutral         c. Against it

2.  Do you believe that women’s rights movements are important?

    1.  Yes, they are important      b. neutral       c. No, they are not important at all

3.  Have you ever been part in women’s rights movement or any related organizations?

    1.  Yes                 b. no, but intended to do so        c. No, not interested at all

4.  Do you believe that having access to abortion is a part of women’s rights?

    1.  Yes                 b. Not sure              c. No

5.  What is your opinion on abortion?

    1.  Support it - her body her choice        b. neutral - it depends     c. against it

6.  If you answer ‘support it’ to the above question, do you believe that women are free to receive abortion even if it’s after 10-12 weeks (which could be medically risky)?

    1.  Yes                b. Not sure               c. No

7.  Have you or someone you know had experienced abortion?

    1.  Yes, myself        b. yes, someone I know      c. No, I don’t know anyone who had experienced abortion

8.  Do you believe that victims of rape shuold be free to receive abortion?

    1.  Yes       b. No

9.  Do you believe that married women should also be free to receive abortion?

    1.  Yes       b. No

10. In your opinion, should abortion always be legal?

    1.  Yes       b. No

# References
