---
title: "Changing Social Perceptions about Women's Rights and Abortion" 
subtitle: "Analysis using US General Social Survey Data (1972 to 2022)"
author: 
  - Kuiyao Qiao
  - Xiaoyu Miao
thanks: "Code and data are available at: https://github.com/QPP123/Investigating-Social-Perceptions-of-Women-s-Rights-and-Abortion-Over-Years-A-Research-with-Surveys"
date: "March 11, 2024"
date-format: long
abstract: "We investigate changes in the social perception of abortion in the United States over the past five decades, utilizing data from the General Social Survey. We analyse 46,700 survey responses and find that opinions on abortion have slightly shifted over time. Support for abortion depends on aspects such as the health of the woman, economic constraints, marital status, fetal health issues, and whether there was rape. This study highlights the critical role of social perception in shaping abortion-related legislation, healthcare access, and the broader discourse on gender equality and women's autonomy."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#download the packages if necessary, then load the packages
library(readxl)
library(dplyr)
library(ggplot2)
library(here)
library(readr)

file_path <- here("inputs", "data", "analysis_data", "analysis_data.csv")
data <- read_csv(file_path)

```

# Introduction

Abortion is an essential yet controversial topic that is worth studying as it relates to lives. We are motivated to write this topic because society's general perceptions of women's rights and abortion can significantly affect numerous women's fates – whether they could be financially self-sufficient or not, whether they get to choose to become a mother or not, whether they can live their life according to their free will or not. As half of the world's population is biologically female, and almost everyone has female family members and female friends, the topic is directly related to nearly everyone. This paper analyzes U.S. social perceptions of abortion from 1972 to 2022, using General Social Survey data. It examines public opinion on various scenarios where abortion might be sought, revealing slight changes towards acceptance in specific contexts, like health risks and rape. The study showcases nuanced views of abortion, indicating evolving social norms. Through detailed analysis, including statistical graphs and demographic insights, it contributes to discussions on women's rights and policy-making, highlighting the complex relationship between societal attitudes and abortion.

In the current academic world, the studies of social perception of abortion have not been conducted in a comparative and historical context. Therefore, we intend to investigate the gap between the lack of research on abortion in a historical context in a holistic approach by sourcing and analyzing the data from a relatively reliable website. We gained the data from GSS (General Social Survey), a sociological survey created by the National Opinion Research at the University of Chicago. This ensures the data source is relatively reliable, actively accurate, and up-to-date. We downloaded the survey data and made graphs and tables out of them, and then we used those to analyze the patterns. 

The paper finds that social perceptions of abortion in the U.S. from 1972 to 2022 have shown slight but not significant changes, with a general trend towards acceptance in specific contexts such as health risks and rape. It reflects the complexity of societal attitudes and the impact of various factors on these opinions. The analysis suggests evolving social norms and points towards the need for further research on demographic specifics to understand these changes better. These findings are significant because the social perception of abortion significantly impacts lawmaking, healthcare access, and individual experiences, as public opinion shapes abortion-related legislation, influences the availability of services, and affects the stigmatization or support individuals face. Moreover, these perceptions guide educational campaigns on reproductive health and underscore the connection between abortion views and broader issues of gender equality and women's autonomy. This complex interplay between social attitudes and practical outcomes highlights the importance of understanding and addressing public opinions on abortion. Seeing the historical trends through survey answers in the past five decades could tell us about social perceptions on abortion over the decades and thus lead us to understand social movements' development better.

Structured into six parts—abstract, introduction, data, results, discussion, and supplementary survey questions—this paper endeavors to provide a deep understanding of the evolution of abortion perceptions in the U.S., shedding light on the societal dynamics that influence these views and their implications for policy and individual experiences. In addition, this paper draws from related literature to dive deeper into the branches of this topic.

Estimand: This study estimates the nation's perception based on the valid survey responses of around 46700 cases. Although the total cases are around 72000, only 46700 instances are valid. Compared to the whole population of the United States, this quantity/parameter is not significant. However, it does represent a general sense of social perceptions. With this being said, we do want to recognize the limitation of this survey study: the demography of the respondents might be limited to certain regions/genders/age ranges/socioeconomic statuses, which could impact the inclusiveness and accuracy of the results.

# Data

This paper uses data from the U.S. General Social Survey (GSS) conducted by NORC at the University of Chicago, where public opinion research on abortion in various contexts provides a window into societal values and moral standards. The GSS data offer a rich source of information that captures different perspectives on the different situations in which abortion might be considered. For this analysis, we focus on a set of survey questions probing respondents' views on whether abortion should be legal under various circumstances.

## Source data

The primary data comes from the GSS cumulative data file (1972-2022), a cross-sectional data set combining all surveys over these 50 years. We use the variable "id" in the data set to rank respondents. One variable is "year". The purpose of this variable is to collect the year in which the survey was conducted. Another variable is the respondents' answers, meaning that when they see the question, they will answer "YES," "NO," or any other answer. From this rich data set, we extracted seven key variables gauging attitudes toward abortion access:

abdefect: Abortion if there is a strong chance of serious defect in the baby abnomore: Abortion if married and doesn't want more children abhlth: Abortion if a woman's health is seriously endangered abpoor: Abortion if low income and can't afford more children abrape: Abortion if pregnancy is the result of rape absingle: Abortion if not married and doesn't want to marry a man abany: Abortion for any reason For each question, respondents could answer "Yes," "No," or "Don't know." "Don't know" responses were treated as missing data to facilitate analysis.

## Data Limitations

### Gap in 1986 data

We're missing the data from the 1986 General Social Survey (GSS), and it's a big deal because it leaves a hole in the data set. The GSS usually gives us a year-by-year look at what people think and feel, so when a year is missing, it messes up the flow and makes it challenging to compare how things change over time. With the 1986 data, it's easier to spot the small changes in what people thought back in the mid-80s. Also, we might be overlooking some significant events or changes in people's opinions from that year, which could help us understand better how things have evolved since then.

### "NA" responses

Dealing with a lot of "don't know" answers in the General Social Survey (GSS) is a pain because it leaves us hanging without clear "yes" or "no" replies, especially when looking at views on abortion. These "NA" or "don't know" choices pop up if people dodge the question or don't have a strong opinion, which punches holes in our data. It's easier to analyze things properly when we need picture chunks. If we see a bunch of these "don't know" answers in certain areas or times, it might mean folks are confused, not well-informed, or don't want to say what they think about abortion. This makes it easier to figure out what the majority believes. These "don't know" replies can shrink the amount of data we have to work with since they're often tossed out like they're not even there, messing with the study's accuracy and how well its findings apply to everyone. So, when we talk about the GSS and abortion opinions, we've got to consider how these "NA" responses limit what we can learn from the data. We need the data from the 1986 General Social Survey (GSS), and it's a big deal because it leaves a hole in the data set. The GSS usually gives us a year-by-year look at what people think and feel, so when a year is missing, it messes up the flow and makes it challenging to compare how things change over time. With the 1986 data, it's easier to spot the small changes in what people thought back in the mid-80s. Also, we might be overlooking some significant events or changes in people's opinions from that year, which could help us understand better how things have evolved since then.

## Data Cleaning

The data was cleaned and analyzed using the open-source statistical programming language R \@\[citeR\]. It uses functionalists from readxl, dplyr, ggplot2, here, and readr. The cleaning process focuses on transforming the data from the third column onward. It uses the mutate and across functions in dplyr to standardize the responses by converting any values that are not "yes" or "no" to NA (missing values), ensuring that the data set contains only clear binary responses for analysis. This step is critical to maintaining the consistency and reliability of the data set, as it filters out ambiguous or invalid entries. Finally, the cleaned data is saved as a CSV file in the specified "analysis_data" directory. This allows access to the cleaned data set for further analysis or reporting. Using write.csv with line name = FALSE ensures the output file is neat.

# Results

Results will likely require summary statistics, tables, graphs, images, and possibly statistical analysis or maps. There should also be text associated with all these aspects. Show the reader the results by plotting them where possible. Talk about them. Explain them. That said, this section should strictly relay results. Regression tables must not contain stars.

**Graph 1:**

Key points are:

-   The 'Yes' responses (green line) are consistently higher than the 'No' responses (red line) from 1970 to slightly past 2020, indicating a prevailing agreement with the option of abortion under these circumstances.

-   There's a noticeable spike in 'Yes' responses around the year 2000, followed by a sharp decline and then another rise. These fluctuations could be due to a variety of factors, including shifts in public awareness, policy changes, or external events influencing public opinion.

-   The 'No' responses show a slight increase over time but remain relatively low compared to the 'Yes' responses, suggesting that opposition to abortion in the case of potential serious defects is less common.

For this research, these trends could highlight that there is generally strong support for the availability of abortion when there's a risk of serious defects in the baby, but it's important to consider the context behind the fluctuations in responses for a comprehensive analysis.

```{r}
#| label: a strong chance of serious defect in the baby
#| fig-cap: "a strong chance of serious defect in the baby"
#| fig-subcap: "[@NORC_2023a]"
#| echo: false
#| warning: false
#| message: false

filtered_data_abdefect <- data %>%
  filter(abdefect %in% c("YES", "NO"))

data_summary_abdefect <- filtered_data_abdefect %>%
  group_by(year, abdefect) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abdefect, aes(x = year, y = count, color = abdefect, group = abdefect)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

**Graph 2:**

Observations from the graph:

-   The 'Yes' responses (green) and 'No' responses (red) are quite volatile over the years. There are times when 'Yes' is higher and others when 'No' dominates, suggesting fluctuating opinions on this issue.

-   Around 2000 and just after 2010, there are sharp spikes in both 'Yes' and 'No' responses, which could be tied to specific events or changes in societal attitudes at those times.

-   Despite the fluctuations, 'Yes' responses tend to be higher on average in the later years, especially past 2010, hinting at a possible trend towards greater acceptance of abortion under these circumstances.

In this research, these trends could point to evolving social norms around family planning and autonomy within marriage, with a more recent tilt towards supporting a married woman's choice not to have more children. The variability could also reflect the impact of policy changes or public debates occurring at different times.

```{r}
#| label: she is married and does not want any more children
#| fig-cap: she is married and does not want any more children
#| fig-subcap: "[@NORC_2023b]"
#| echo: false
#| warning: false
#| message: false

filtered_data_abnomore <- data %>%
  filter(abnomore %in% c("YES", "NO"))

data_summary_abnomore <- filtered_data_abnomore %>%
  group_by(year, abnomore) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abnomore, aes(x = year, y = count, color = abnomore, group = abnomore)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

**Graph 3:**

Here's what the data indicates:

-   The 'Yes' responses (green) are substantially higher than the 'No' responses (red) throughout the entire period from the 1970s to 2020. This indicates strong and consistent support for a woman's right to an abortion when her health is at risk.

-   The 'Yes' line shows some fluctuation over the years, with peaks and valleys, but the overall support remains high.

-   The 'No' responses are relatively flat and low in comparison, suggesting that there is minimal opposition to abortion under these circumstances and that this view has not changed dramatically over time.

In the context of this research, this graph could suggest that there is a general consensus supporting the necessity of abortion when it comes to health concerns, and this has been a relatively stable opinion over the last five decades.

```{r}
#| label: womans own health is seriously endangered by the pregnancy
#| fig-cap: womans own health is seriously endangered by the pregnancy
#| fig-subcap: "[@NORC_2023c]"
#| echo: false
#| warning: false
#| message: false

filtered_data_abhlth <- data %>%
  filter(abhlth %in% c("YES", "NO"))

data_summary_abhlth <- filtered_data_abhlth %>%
  group_by(year, abhlth) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abhlth, aes(x = year, y = count, color = abhlth, group = abhlth)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

**Graph 4:**

Key observations are:

-   'Yes' responses (green) generally exceed 'No' responses (red), indicating more support for allowing abortions under these financial circumstances.

-   The data is quite volatile, with many peaks and troughs, suggesting that public opinion on this issue has varied significantly over time.

-   Around the late 1990s and again in the late 2000s, there are moments where 'No' responses surpass 'Yes' responses, which could be indicative of particular events or changes in societal attitudes that influenced public opinion during those times.

This data is valuable for this research as it demonstrates that economic factors affecting decisions on abortion have been a point of contention over the years, with no clear consensus emerging. It might be useful to investigate the societal and economic contexts corresponding to the shifts in public opinion reflected in the graph.

```{r}
#| label: family has a very low income and cannot afford any more children
#| fig-cap: family has a very low income and cannot afford any more children
#| fig-subcap: -"[@NORC_2023d]"
#| echo: false
#| warning: false
#| message: false

filtered_data_abpoor <- data %>%
  filter(abpoor %in% c("YES", "NO"))

data_summary_abpoor <- filtered_data_abpoor %>%
  group_by(year, abpoor) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abpoor, aes(x = year, y = count, color = abpoor, group = abpoor)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

**Graph 5:**

Observations include:

-   'Yes' responses (green) are significantly higher than 'No' responses (red) throughout the entire time span, indicating strong support for allowing abortion under these circumstances.

-   There are notable fluctuations in the 'Yes' responses, with peaks and valleys, but the overall trend remains much higher than the 'No' responses.

-   The 'No' responses are relatively consistent over the years and remain low, suggesting that opposition to abortion in this context is consistently minimal.

From this data, one could infer that there is a stable and overwhelming support for a woman's right to choose abortion in the case of rape, which has been maintained over the past five decades. This could be a key point in your research, reflecting a societal consensus on the right to abortion in such traumatic circumstances.

```{r}
#| label: she became pregnant as a result of rape
#| fig-cap: she became pregnant as a result of rape
#| fig-subcap: -"[@NORC_2023e]"
#| echo: false
#| warning: false
#| message: false

filtered_data_abrape <- data %>%
  filter(abrape %in% c("YES", "NO"))

data_summary_abrape <- filtered_data_abrape %>%
  group_by(year, abrape) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_abrape, aes(x = year, y = count, color = abrape, group = abrape)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

**Graph 6:**

Observations include:

-   Unlike the previous graph, here, the 'Yes' (green) and 'No' (red) responses show a more competitive trend, with 'No' responses occasionally matching or exceeding 'Yes' responses at several points in time.

-   The 'Yes' responses show high volatility but maintain an edge over 'No' responses for a majority of the time period.

-   The late 1990s through the 2000s show particularly sharp fluctuations, with 'No' responses peaking above 'Yes' responses more frequently compared to other decades.

This data suggests that public opinion on abortion in the case of rape has been more divided at certain times, with significant changes in attitude that may correlate with cultural, legal, or political events. It could be valuable to correlate these shifts with specific historical contexts for a deeper understanding in your research.

```{r}
#| label: she is not married and does not want to marry the man
#| fig-cap: she is not married and does not want to marry the man
#| fig-subcap: -"[@NORC_2023f]"
#| echo: false
#| warning: false
#| message: false

filtered_data <- data %>%
  filter(absingle %in% c("YES", "NO"))

data_summary_absingle <- filtered_data %>%
  group_by(year, absingle) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(data_summary_absingle, aes(x = year, y = count, color = absingle, group = absingle)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

**Graph 7:**

Notable points:

-   There's a close contention between 'Yes' (green) and 'No' (red) responses throughout the years, reflecting a more divided public opinion on the broader issue of abortion on demand.

-   The 'No' responses have peaks that often surpass the 'Yes' responses, particularly in the mid-1980s to the mid-1990s and again in the early 2000s.

-   Both responses exhibit significant volatility, which could be indicative of shifting social, political, and legal landscapes influencing public opinion on the issue.

This graph suggests that when considering abortion without condition, public opinion has been more polarized, with shifts possibly aligned with broader debates and changes in societal attitudes towards abortion.

```{r}
#| label: abortion if woman wants for any reason
#| fig-cap: abortion if woman wants for any reason
#| fig-subcap: "[@NORC_2023g]"
#| echo: false
#| warning: false
#| message: false

data_summary_abany <- data %>%
  filter(abany %in% c("YES", "NO")) %>%
  group_by(year, abany) %>%
  summarise(count = n(), .groups = 'drop') %>%
  mutate(abany = factor(abany, levels = c("YES", "NO"))) # Ensure consistent ordering for the line plot

ggplot(data_summary_abany, aes(x = year, y = count, color = abany, group = abany)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  labs(x = "Year",
       y = "Count of Responses",
       color = "Response") +
  scale_color_manual(values = c("YES" = "darkgreen", "NO" = "red"))
```

The seven graphs you provided show public opinion over time on various circumstances under which abortion might be sought. Here's a consolidated summary of insights from the data:

-   Health of the woman: There is consistent and strong support for abortion when the woman's health is seriously endangered. This opinion seems to hold relatively steady over the decades.

-   Economic reasons: Opinions on abortion due to economic reasons, such as the family being of very low income and unable to afford more children, show more fluctuation and suggest that this reason for abortion is more contentious and less consistently supported than health-related reasons.

-   Marital status: There appears to be more support for abortion rights for unmarried women compared to married women who do not want more children, although both show some level of volatility in responses over the years.

-   Fetal health issues: There is strong support for abortion in cases of a strong chance of a serious defect in the baby, but with some fluctuations that might be tied to external factors or events.

-   Pregnancy resulting from rape: Support for abortion is very high when pregnancy is the result of rape, with minimal opposition across the decades.

-   General abortion access: When considering abortion for any reason the woman wants, the data shows a divided opinion with significant contention and volatility over the years.

From these graphs, we can find that public opinion on abortion is generally more supportive when the reasons are related to the health of the woman or the fetus, or in cases of rape. When the reasons for seeking an abortion are not tied to such specific conditions, like economic reasons or general access to abortion for any reason, the public opinion seems to be more divided and subject to change over time.

The volatility in responses, particularly in the graphs related to economic reasons and general abortion access, suggests that these topics are more influenced by the social, political, and economic context at various points in time. These insights could be critical for your research as they demonstrate the complexity and nuance of public opinion on abortion. It would be essential to analyze these trends in relation to policy changes, economic conditions, and social attitudes to understand the underlying causes of the fluctuations in opinion.

# Discussion

This paper examines the social perception of under what circumstances women should be free to receive an abortion. The underlying theme is the autonomy of women to their bodies:

Not only did we learn more about the social perception of Abortion over five decades in the United States, Overall, the social perception of women receiving abortions because of a strong chance of severe defects in the baby has not changed significantly over the decades;

However, holistically in recent ten years, more and more people support married women who don't want any more children to receive Abortion (figure 2); more and more people support women to be accessible to receive Abortion if the family is not financially capable of affording any more children (figure 4); more and more people support Abortion if the women get pregnant as a result of rape; more and more people support Abortion if women want for any reason. All these show that the U.S. society is progressing towards a more inclusive and liberal environment for women.

**Strengths:**

We found out a lot about what people in the U.S. think about abortion over the last fifty years. One of the big takeaways is that people's views on abortion haven't changed much when it comes to severe health defects in a baby. But in the last decade, we've seen more support for abortion in cases like when a married woman doesn't want more children when families can't afford more kids when pregnancy is due to rape, or even if a woman wants an abortion for any reason. This shows that America is slowly becoming more open and supportive of woman's rights to choose abortion.

**Weaknesses:**

The study doesn't tell us how they gathered all this info, making me wonder who they asked and how they asked it. Were they all from a particular age group or background? That could make the results biased. The survey also seems too broad and doesn't dive deep into specific details like age, race, gender, or where people live. These things matter because they can change how people view abortion. Plus, politics, religion, and culture play a massive role in shaping these views. Still, the study doesn't fully connect the dots on how all these factors interact with people's opinions on abortion.

The whole survey is too general from an academic point of view. Age? Race? Gender? Religious? Geography? Perceptions are different

The phenomena this research reveals are complex due to many factors, including politics, religion, socio-cultural progressions, and economic changes. According to related scholarly literature, the article "Introduction: The Politics of Abortion 50 Years after Roe" by Katrina Kimport and Rebecca Kreitzer reveals several critical insights about the intricate relationship between Abortion and politics in the United States. The article touches on the topic of Abortion and social perception, particularly highlighting how Abortion has been both marginalized and siloed in social science research despite its significance in politics and its common occurrence as a pregnancy outcome. This marginalization reflects broader social perceptions and the contentious nature of Abortion as a public issue. The authors discuss the "abortion research paradox," where, despite the political and social importance of Abortion, it remains underrepresented in top disciplinary journals, suggesting a disconnect between its real-world significance and its academic treatment.

To delve deeper into how different races, ethnicity, and cultures view Abortion, we accessed the scholarly article from Front Public Health named 'The role of community and culture in abortion perceptions, decisions, and experiences among Asian Americans'. This study explores the influence of culture and community on Asian Americans' experiences with medication abortion in the U.S. It suggests that religious stigma and lack of openness about sexual and reproductive health, among other factors, contribute to the complexities Asian Americans face regarding Abortion. It emphasizes the need for culturally sensitive and inclusive abortion care to address these challenges. It also advocates the need for more inclusive abortion care services that consider the diverse backgrounds and experiences within the Asian American community.

To understand how different religions view Abortion, the Pew Research Center's study on views about Abortion reveals varying opinions across different religious groups. Buddhists and Jews show high support for legal Abortion in most/all cases, while Evangelical Protestants and Jehovah's Witnesses exhibit strong opposition. Catholics and Mainline Protestants are more divided. The study highlights the complex relationship between religious beliefs and abortion attitudes, underscoring significant differences even within broader religious traditions. Detailed Statistics are provided in the article.

What is also worth mentioning is the legal progression of Abortion in the United States. Roe v. Wade was a landmark 1973 Supreme Court decision that recognized the constitutional right to Abortion. This decision required laws against Abortion to be balanced against the woman's right to privacy, significantly changing abortion laws across the United States and sparking ongoing debates and legislative actions around abortion rights. However, the government policy called the Global Gag Rule in 1984 came out to essentially go against Abortion. It prohibited foreign NGOs receiving U.S. federal funding from providing, advocating for, or referring patients to abortion services, even with non-U.S. funds. It has been rescinded and reinstated several times depending on the administration in power, reflecting changing U.S. political attitudes towards Abortion. This policy significantly impacts global health organizations' ability to offer comprehensive reproductive health services. In 1992, Planned Parenthood v. Casey appeared to become a significant U.S. Supreme Court case that reaffirmed the constitutional right to Abortion established in Roe v. Wade (1973). However, it also allowed states more leeway to impose restrictions on Abortion, introducing the "undue burden" standard. This standard prohibits regulations that place a substantial obstacle in the path of a woman seeking an abortion before fetal viability. Finally, in June 2022, Roe v. Wade was overturned by the U.S. Supreme Court on June 24, 2022, removing the federal constitutional right to an abortion that had been in place for nearly 50 years. All of these historical events represent the complex history of abortion rights in the United States. We can see that in the 50 years (1972-2022), the historical trend of Abortion at a legal level in the U.S. was not progressing in a single direction. It was established, restricted, prohibited, liberated, and overturned. This reflects its complicated relationship to U.S. politics – the legislative decisions heavily depend on the views of the politicians at the time. This also contributes to our understanding of why the social perceptions varied through different times in the data.

How can we proceed in the future:

In the future, the data would benefit if they specify survey respondents' demographic information such as age, race, gender, religion, and geography. From the research above from scholarly literature, we can understand how politics, cultures and beliefs influence the social perception of Abortion.

Discussion of the questionnaire: The good part is that each question is particular and on-point: they are clearly yes or no questions but also leave the respondents the option of 'don't know' or 'not applicable,' which is professionally designed. The bad part of the questionnaire is that the questions did not cover all the scenarios: for example, whether or not to receive an abortion could be considered very differently if the pregnancy is over 10 or 12 weeks. The survey questions should add questions regarding this, such as 'Do you support abortion if the pregnancy is over ten weeks (which could be medically risky)?' or something like that.

\newpage

# Appendix

**Survey**

The survey should have an introductory section and include the details of a contact person. The survey questions should be well constructed and appropriate to the task. The questions should have an appropriate ordering. A final section should thank the respondent.

**Supplementary Survey**

Introduction to the survey: Women’s rights is an essential topic in nowadays society. Within this topic, abortion is a controversial topic. The General Social Survey collects information on people’s opinions and perceptions on those topics. In order to deepen our understandings and analysis of the topics, we designed this survey to collect people’s opinions and perceptions of women’s rights and abortion. We conduct the research studies with good faiths.

Thank you for all the respondents for completing this survey! We appreciate your time and effort. Your response will help us gain a deeper understanding of people’s perceptions on women’s rights and abortion.

Supplementary Survey Questions:

1.  What is your opinion on women’s rights?

    1.  Support it           b. Neutral         c. Against it

2.  Do you believe that women’s rights movements are important?

    1.  Yes, they are important      b. neutral       c. No, they are not important at all

3.  Have you ever been part in women’s rights movement or any related organizations?

    1.  Yes                 b. no, but intended to do so        c. No, not interested at all

4.  Do you believe that having access to abortion is a part of women’s rights?

    1.  Yes                 b. Not sure              c. No

5.  What is your opinion on abortion?

    1.  Support it - her body her choice        b. neutral - it depends     c. against it

6.  If you answer ‘support it’ to the above question, do you believe that women are free to receive abortion even if it’s after 10-12 weeks (which could be medically risky)?

    1.  Yes                b. Not sure               c. No

7.  Have you or someone you know had experienced abortion?

    1.  Yes, myself        b. yes, someone I know      c. No, I don’t know anyone who had experienced abortion

8.  Do you believe that victims of rape shuold be free to receive abortion?

    1.  Yes       b. No

9.  Do you believe that married women should also be free to receive abortion?

    1.  Yes       b. No

10. In your opinion, should abortion always be legal?

    1.  Yes       b. No

# References
